---
title: "Assignment 5"
author: "Kai Lukowiak"
date: '2018-03-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Choose independently two numbers B and C at random from the interval [0, 1] with uniform density. Prove that B and C are proper probability distributions. Note that the point (B,C) is then chosen at random in the unit square. Find the probability that


```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(ggthemes)
set.seed(101)
B <- runif(10000, 0, 1)
C <- runif(10000, 0, 1)

sumVec <- B + C
productVec <- B * C
absVec <- abs(B - C)
maxVec <- pmax(B, C)
minVec <- pmin(B, C)

df <- data.frame(cbind(B, C))
ggplot(df) + 
  geom_histogram(aes(x = df$B), fill = 'grey', alpha = .5, color = 'black')+
  geom_histogram(aes(df$C), fill = 'light blue', alpha = .5, color = 'black') +
  ggtitle('Uniform Density Distrubution') + theme_tufte()
```

# a)
$B + C < 1/2$
Since B + C is a straight line, the function to integrate over is simply $x$.
```{r}
IntagreteBootStrap <- function(vec, integrand){
  ans1 = integrate(integrand, lower = 0, upper = 1/2)$value
  ans2 = ecdf(vec)(.5)
  print(paste('The answer found from integratin is:', ans1, 'The answer found from boot strap is:', ans2))
}
integrand = function(x){x}
IntagreteBootStrap(sumVec, integrand)
```


# b)
$BC < 1/2$
$B\cdot C$ 


```{r}
integrand = function(x){-log(x)}
IntagreteBootStrap(productVec, integrand)
```


# c)

$|B − C| < 1/2$


```{r}
integrand = function(x){2 * (1 - x)} # multiply by 2 because of the positive and neg options.
IntagreteBootStrap(absVec, integrand)
```


# d)
$max(B,C) < 1/2$

Instead of integrating, we can use more elementary probability theory to solve this. The probability of  $max(B,C) < 1/2$ is given by:

$$ P(B) < 1/2 \cap P(C) < 1/2$$
Since $E(B,C) = \frac{1}{2}, \frac{1}{2}$ The solution is $\Big ( \frac{1}{2} \Big ) ^2 = \frac{1}{4}$ 
```{r}
length(maxVec[maxVec<.5]) / length(maxVec)
```



# e)
$min(B,C) < 1/2$

Since $E(B,C) = \frac{1}{2}, \frac{1}{2}$ The solution is $1 - \Big ( \frac{1}{2} \Big ) ^2 = \frac{3}{4}$

```{r}
length(minVec[minVec<.5]) / length(minVec)
```

