---
title: "Assignment 5"
author: "Kai Lukowiak"
date: '2018-03-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Choose independently two numbers B and C at random from the interval [0, 1] with uniform density. Prove that B and C are proper probability distributions. Note that the point (B,C) is then chosen at random in the unit square. Find the probability that


```{r warning=FALSE, message=FALSE}
set.seed(101)
A <- runif(10000, 0, 1)
B <- runif(10000, 0, 1)
df <- data.frame(A = A, B = B)
library(tidyverse)


ggplot(df) + 
  geom_histogram(aes(x = df$A), fill = 'red', alpha = .5)+
  geom_histogram(aes(df$B), fill = 'blue', alpha = .5) +
  ggtitle('Uniform Density Distrubution')

```

# a)
$B + C < 1/2$

```{r }
stat1 <- df %>% 
  mutate(sumAB = A + B) %>% 
  filter(sumAB < .5) %>% 
  select(sumAB) %>% 
  summarise(ans <- n()/nrow(df))
statP <- stat1$`ans <- n()/nrow(df)`

```

The chance of the sum of  any two entries (A, B) being less than one half is `r statP * 100`%


# b)
$BC < 1/2$


```{r}
stat1 <- df %>% 
  mutate(sumAB = A * B) %>% 
  filter(sumAB < .5) %>% 
  select(sumAB) %>% 
  summarise(ans <- n()/nrow(df))
statP <- round(stat1$`ans <- n()/nrow(df)`,2)

```

The chance of the product of any two entries of (A, B) being less than one half is `r statP * 100`%



# c)

$|B − C| < 1/2$

```{r}
stat1 <- df %>% 
  mutate(sumAB = abs(A - B)) %>% 
  filter(sumAB < .5) %>% 
  select(sumAB) %>% 
  summarise(ans <- n()/nrow(df))
statP <- round(stat1$`ans <- n()/nrow(df)`,2)
```

The chance of the subtraction of any two entries of (A, B) being less than one half is `r statP * 100`%

# d)
$max{B,C} < 1/2$

```{r}
stat1 <- df %>% 
  mutate(sumAB = max(A, B)) %>% 
  filter(sumAB < .5) %>% 
  select(sumAB) %>% 
  summarise(ans <- n()/nrow(df))
statP <- round(stat1$`ans <- n()/nrow(df)`,2)
```

The chance of the max value of any two entries of (A, B) being less than one half is `r statP * 100`%

# e)
```{r}
stat1 <- df %>% 
  mutate(sumAB = min(A, B)) #%>% 
  #filter(sumAB < .5) %>% 
  #select(sumAB) %>% 
  #summarise(ans <- n()/nrow(df))
#statP <- round(stat1$`ans <- n()/nrow(df)`,2)
```

The chance of the min value of any two entries of (A, B) being less than one half is `r statP * 100`%
