---
title: "Week13 Discussion"
author: "Kai Lukowiak"
date: '2018-05-02'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 4.2.3
Water flows onto a flat surface at a rate of 5cm3/s forming a
circular puddle 10mm deep. How fast is the radius growing
when the radius is:
a) 1 cm
b) 10 cm
c) 100 cm

Let $V=volume$, $r=radius$, $A=Area$ and $d = depth$.


$$
V=A \cdot d
\\
\therefore
\\
V=A \cdot 0.1cm
\\
\&
\\
\frac{\partial}{\partial t}= 0.1cm \frac{\partial A}{\partial t}
$$
We know that $\frac{\partial V}{\partial t} = 5cm \quad \therefore \quad 5=0.1 \frac{\partial A}{\partial t}$ 
therefor $\frac{\partial A}{\partial t} = 50$


$$
A=\pi r^2
\\
\frac{\partial A}{\partial t}=\frac{\partial}{\partial t}(\pi r^2)
\\
\frac{\partial A}{\partial t}=2 \pi r \frac{\partial r}{\partial t}
$$

Since we know that $\frac{\partial A}{\partial t} = 50$ then $50 = 2 \pi r \frac{\partial r}{\partial t}$
Thus the rate of change is.
$$
\frac{50}{ 2 \pi r } = \frac{\partial r}{\partial t}
$$

Thus to answer our questions:

```{r}
nums <- c(1,10,100)

ans <- function(r){
  50/(2 * pi * r)
}

ans(nums)
```

Are the respective rates of change.
$$
\delta
$$