---
title: "HomeWorkW1"
author: "Kai Lukowiak"
date: "January 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem Set 1
## 1
Calculate the dot product u.v where u = [0.5; 0.5] and v = [3; −4]

```{r}
u <- c(0.5, 0.5)
v <- c(3, -4)
print(paste("The dot product of u, v is:", u %*% v))
```

## 2
What are the lengths of u and v? Please note that the mathematical notion of the
length of a vector is not the same as a computer science definition.

The length of a vector is given by the euclidean distance or by use of the Pythagoras therm. 

The formula for this is $x^2 + y^2 = z^2$ as such:

```{r}
LenU <- sqrt(sum(u^2))
LenV <- sqrt(sum(v^2))
print(paste('The length of u is ', LenU, 'and the length of v is', LenV))
```

## 3
What is the linear combination: 3u − 2v? 
```{r}
linComb <- 3 * u - 2 * v
print(paste(c('The linear combination of the vectors and constants is:', linComb), collapse = " "))
```

## 4 

What is the angle between u and v?

The angel can be given with the $cosin$ distance.

Special thanks to [this stackoverflow question.](https://stackoverflow.com/questions/1897704/angle-between-two-vectors-in-r)

```{r}
theta <- acos( sum(u*v) / ( sqrt(sum(u * u)) * sqrt(sum(v * v)))) 
theta * 180 / pi
```

# Problem Set 2

Set up a system of equations with 3 variables and 3 constraints and solve for x. Please
write a function in R that will take two variables (matrix A & constraint vector b) and
solve using elimination. Your function should produce the right answer for the system of
equations for any 3-variable, 3-equation system. You don’t have to worry about degenerate
cases and can safely assume that the function will only be tested with a system of equations
that has a solution. Please note that you do have to worry about zero pivots, though. Please
note that you should not use the built-in function solve to solve this system or use matrix
inverses. The approach that you should employ is to construct an Upper Triangular Matrix
and then back-substitute to get the solution. Alternatively, you can augment the matrix
A with vector b and jointly apply the Gauss Jordan elimination procedure.
12
IS 605 FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS - FALL 2014
Please test it with the system below and it should produce a solution x = [−1.55, −0.32, 0.95]

$$\mathbf{} \left[\begin{array}
{rrr}
1 & 1 & 3 \\
2 & -1 & 5 \\
-1 & -2 & 4
\end{array}\right]
%
\left[ \begin{array}{ccc}
1  \\
2 \\
6
\end{array} \right]
$$


Please send your code (as an R-markdown file, named using your first initial, last name,
assignment, problem set. For instance, if I submit the code for this assignment, it will be
called GIyengar Assign1.Rmd

## Solution:

Instead of writing out a function that would do the exact steps I wanted to use loops to make my code shorter. I went down quite the rabbit hole trying to make it work with `for` loops but this was exceedingly difficult. 

I stumbled upon [this](https://stackoverflow.com/questions/16044377/how-to-do-gaussian-elimination-in-r-do-not-use-solve) that would work on any n dimensional matrix. I modified it to work in a function. The result as shown below the initial solution works on matrixies of any size.

```{r}

matrixA <- matrix( 
  c(1,1,3,2,-1,5,-1,-2,4),
  nrow=3,
  byrow = TRUE)

vectorb <- matrix( 
  c(1,2,6),
  ncol=1,
  byrow = TRUE)
```

```{r}
loopFunc <- function(A, b){
  p <- nrow(A)
  aug <- cbind(A, b)
  aug[1, ] <- aug[1, ] / aug[1,1] # Normaliz the fist number to 1. This makes everything easier.
  
  i <- 2 # So that we don't run the code on an index out of range e.g., A[1-1, ]
  while (i < p + 1) {
    j <- i # This effectively sets the loop to go over one set and then the other. 
    while (j < p + 1) {
      aug[j, ] <- aug[j, ] - aug[i -1, ] * aug[j, i-1] # The last argument here works because we normalized the first row to 1
      # The above command also sets the jth row of the matrix to the correct result.
      j <- j + 1 # This makes the code easier than in a for loop.
    }
    while (aug[i, i] == 0) {
      aug <- rbind(aug[-1], aug[i]) # Checks to make sure that we don't have zero devision 
    }
    aug[i, ] <- aug[i, ] / aug[i, i] # Scales the row  to 1 again, making calculations easier. 
    i <- i + 1
  }
  for (i in p:2) {
    # p is now a large number and we need to go in reverse to solve the upper triangle.
    for (j in i:2-1) { # The same logic holds here but for the second element. 
      aug[j, ] <- aug[j, ] - aug[i, ] * aug[j, i] # Solves the rows going up. 
    }
  }
  return(aug)
}
mx <- loopFunc(matrixA, vectorb)
mx
```


### N > 3 Dimensions:

```{r}
set.seed(101)
A <- matrix(floor(runif(100, 1,11)) , 10)
b = floor(runif(10, 1 , 11))
A
b
loopFunc(A, b)
```

