---
title: "Assignment 10"
author: "Kai Lukowiak"
date: '2018-04-04'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars.
A guard agrees to make a series of bets with him. If Smith bets A dollars,
he wins A dollars with probability .4 and loses A dollars with probability .6.
Find the probability that he wins 8 dollars before losing all of his money if

# (a) he bets 1 dollar each time (timid strategy).

This is a random walk problem with two absorbing states, 0 and 8. 

From pg. 487 of *Introduction to Probability* by Charles M. Grinstead:

$$
q = 0.6
\\
p = 1-q = 0.4
\\
s = 1
\\
M = 8
\\
P = \frac{1 - \frac{q}{p}^s}{1 - \frac{q}{p}^M}
$$
In R we find:

```{r}
q = 0.6
p = 0.4
s = 1
M = 8
P = (1 - (q/p)^s) / (1 - (q/p)^M)
P
```

(b) he bets, each time, as much as possible but not more than necessary to
bring his fortune up to 8 dollars (bold strategy).
1 to 2 to 4 to 8
Starting at 1, he must win 3 succesive bets (or lose everything). This gives 
us the problem of finding the probability of:

```{r}
0.4 ^ 3
```

(c) Which strategy gives Smith the better chance of getting out of jail?

The bold stratagey would be better. 


If he started off with two dollars:


```{r}
B = 0.4 ^2
B
q = 0.6
p = 0.4
s = 2
M = 8
T = (1 - (q/p)^s) / (1 - (q/p)^M)
T
```

So he is still better than if he did the timid stratagy.

This makes sense because the law of large numbers will lead to achieving the expected value (0).

