---
title: "Week 15"
author: "Kai Lukowiak"
date: '2018-05-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1
Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary.

( 5.6, 8.8 ), ( 6.3, 12.4 ), ( 7, 14.8 ), ( 7.7, 18.2 ), ( 8.4, 20.8 )

```{r message=F, warning=F}
library(tidyverse)
library(ggthemes)
X <- c(5.6, 6.3, 7, 7.7, 8.4)
y = c(8.8, 12.4, 14.8, 18.2, 20.8)
z <- rep(0, 5)
df = data_frame(X, y, z)
```


```{r}

ggplot(data = df, aes(x=X, y=y)) +
  geom_point() + 
  theme_tufte() +
  ggtitle('Scatter Plot')
```

```{r}
mod = lm(y ~ X, data = df)
sum1 <- summary(mod)
ans1 <- sum1$coefficients[1]
```

The answer is `r ans1`.


#2
Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form
( x, y, z ). Separate multiple points with a comma.

$$
f(x,y) = 24x-6xy^2 -8y^3
\\
\frac{\partial f}{\partial x}=24 - 6y^2
\\
\frac{\partial f}{\partial y}=-12xy-24y^2
$$

Now we must find where both partials are equal to zero. For x:

$$
24-6y^2=0
\\
4-y^2=0
\\
y^2=4
\\
y = (2, \quad -2)
$$

For y:

$$
-12xy-24y^2=0
\\
xy-2y^2=0 
\\
xy=2y^2
\\
x=2y
$$

Now plug in the values from the partial WRT x:

$$
y=(2,-2)
\\
x=2y
\\
x=(4,-4)
$$

Thus the cirtical points (inflection) points are:

$$
(-4,2) \quad \& \quad (4,-2)
$$

# 3
A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she
sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell $81- 21x + 17y$ units of the "house" brand and $40 + 11x - 23y$ units of the "name" brand.

Find the revenu function and the revenue at x=2.3 and y = 4.1
First we need to multiply each equation by the variable that it represents:

$$
R(x,y) = (81- 21x + 17y)x + (40 + 11x - 23y)y
\\
= 81x- 21x^2 + 17yx + 40y + 11xy - 23y^2
\\
= 81x - 21x^2 + 28yx + 40y  - 23y^2
\\

$$

```{r}
x = 2.3
y = 4.1
81 * x - 21* x^2 + 28 *y*x + 40*y  - 23*y^2
```

# 4
A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a
product each week. The total weekly cost is given by $C(x, y) = \frac{1}{6} x^2 + \frac{1}{6} y^2 + 7x + 25y + 700$, where x is the number of
units produced in Los Angeles and y is the number of units produced in Denver. How many units should be produced in
each plant to minimize the total weekly cost?

This can be written as a univariate function because $x=96-y$

$$
C(x, y) = \frac{1}{6} x^2 + \frac{1}{6} y^2 + 7x + 25y + 700
\\
C(96-y, y) = \frac{1}{6} (96-y)^2 + \frac{1}{6} y^2 + 7(96-y) + 25y + 700
\\
= \frac{1}{6}(9216 - 192y + y^2) + \frac{1}{6} y^2 + 7(96-y) + 25y + 700
\\
= \frac{1}{6}y^2 -32y + 1536 + \frac{1}{6} y^2 + 672-7y + 25y + 700
\\
C(y)= \frac{1}{3}y^2 -14y +2908
\\
C'(y) = \frac{2}{3}y - 14

$$


Costs are minimized when $C'(y)=0$

$$
\frac{2}{3}y - 14=0
\\
 \frac{2}{3}y =14
 \\
 y =21
 \\
 \therefore
 \\
 x = 96-21=75
$$

# 5
Evaluate the double integral on the given region.
$$
\int \int (e^{8x+3y}dA);\quad R:2 \leq x \leq 4 \quad \&\quad \leq y \leq 4
$$

Write your answer in exact form without decimals.

$$
\int_2 ^4 \int_2 ^4 (e^{8x+3y}dx dy)
\\
= \int_2 ^4 \frac{1}{3}e^{8x+3y}|_2^4 dx
\\
= \int_2 ^4 (\frac{1}{3}e^{8x+12}) - (\frac{1}{3}e^{8x+6}) dx
\\
=  (\frac{1}{24}e^{8x+12}) - (\frac{1}{24}e^{8x+6}) |_2 ^4
\\
=  ((\frac{1}{24}e^{44}) - (\frac{1}{24}e^{38})) -  ((\frac{1}{24}e^{28}) - (\frac{1}{24}e^{22}))
\\
=  \frac{1}{24}e^{44} - \frac{1}{24}e^{38} -  \frac{1}{24}e^{28} + \frac{1}{24}e^{22}
\\
 =  \frac{1}{24}(e^{44} - e^{38} - e^{28} + e^{22})
$$