---
title: "Final Exam"
author: "Kai Lukowiak"
date: '2018-05-18'
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```


# Instructions
Your final is due by the end of day on 5/20/2018 You should post your solutions to your GitHub account or RPubs. You are also expected to make a short presentation via YouTube and post that recording to the board. This project will show off your ability to understand the elements of the class.

You are to register for Kaggle.com (free) and compete in the House Prices: Advanced Regression Techniques competition. <https://www.kaggle.com/c/house-prices-advanced-regression-techniques> . I want you to do the following.

# Data Reading and Libraries

```{r}
library(tidyverse)
library(ggthemes)
library(knitr)
library(caret)
library(e1071)
library(glmnet)
```


```{r}
df <- read_csv('train.csv')
df[sapply(df, is.character)] <- lapply(df[sapply(df, is.character)], as.factor)
```

## The data Frame

```{r}
glimpse(df)
```

* Pick one of the quantitative independent variables from the training data set (train.csv) , and define that variable as X. Make sure this variable is skewed to the right!
* Pick the dependent variable and define it as Y


```{r}
library(psych)
skewDF <- describe(df)
test <- skewDF %>% 
  select(skew) %>% 
  mutate(VarName = names(df)) %>% 
  arrange(desc(abs(skew))) %>% 
  head()
test
```

`LotArea` has a large skeww

```{r}
ggplot(data = df, aes(x=LotArea)) + 
  geom_density(colour = ggthemes_data$solarized$base['base00'], 
               fill = ggthemes_data$solarized$accents['blue'], 
               alpha = 0.4) +
  theme_minimal() +
  ggtitle('Lot Area')
```

We can verify this skew with the summary of the data:

```{r}
summary(df$LotArea)
```

This has an obvious right skew.

## Probability
Calculate as a minimum the below probabilities a through c. Assume the small letter "x" is estimated as the 1st quartile of the X variable, and the small letter "y" is estimated as the 1st quartile of the Y variable. Interpret the meaning of all probabilities. In addition, make a table of counts as shown below.


a. $P(X>x | Y>y)$ b. $P(X>x, Y>y)$ c. $P(X<x | Y>y)$


a) $P(B|A) = \frac{P(A) \cup P(B)}{P(A)}$ Therefor, $P(X>x | Y>y) = \frac{P( Y>y) \cup P(X>x)}{P(Y>y)}$

```{r}
# Vars
q1x <- quantile(df$LotArea, 0.25)
q1y <- quantile(df$SalePrice, 0.25)
x <- df %>% filter(LotArea > q1x ) %>% nrow() / nrow(df)
y <- df %>% filter(SalePrice > q1y) %>% nrow() / nrow(df)
xy <- df %>% filter(SalePrice > q1y & LotArea > q1x) %>% nrow()/nrow(df)

yx <- df %>% filter(SalePrice > q1y & LotArea < q1x) %>% nrow()/nrow(df)

a <- (xy)/(y)
b <- x*y
c <- yx/y
```

a) The conditional probability is `r a`

b) The joint probability is `r b`

c) The conditional prbability is `r c`


```{r}
one <- df %>% filter(LotArea <= q1x & SalePrice <= q1y) %>% nrow()
two <- df %>% filter(LotArea > q1x & SalePrice <= q1y) %>% nrow()
three <- one + two
four <- df %>% filter(LotArea <= q1x & SalePrice > q1y) %>% nrow()
five <- df %>% filter(LotArea > q1x & SalePrice > q1y) %>% nrow()
six <- four + five
seven <- one + four
eight <- two + five
nine <- three + six
```

| x/y      | <=1st quartile | >1st quartile | Total |
|----------|----------------|---------------|-------|
| <= 1st q |    `r one`     |    `r two`    |`r three`|
| >1st q   | `r four`       |`r five`       |`r six`|
| Total    |  `r seven`     | `r eight`     |`r nine`|


## Does splitting the training data in this fashion make them independent?



Let A be the new variable counting those observations above the 1st quartile for X, and let B be the new variable counting those observations above the 1st quartile for Y. Does $P(A|B)=P(A)P(B)$? Check mathematically, and then evaluate by running a Chi Square test for association.

```{r}
A <- df %>% filter(LotArea > q1x ) %>% nrow()/ nrow(df)
B <- df %>% filter(SalePrice > q1y ) %>% nrow()/nrow(df)
AB <- df %>% filter(LotArea > q1x, SalePrice > q1y) %>% nrow() / nrow(df)

AB /B
A*B
A
B
```

These results are different meaning that they are not independant because $P(A|B)\neq P(A)$. Thus, splitting them like in the table above will not make them independant.

## Chi Squared Test

```{r}
chisq.test(df$LotArea, df$SalePrice)
```


Since the critical value of $\chi$ is high, we reject the null hypothesis that they are independant. This comes at no surprise since the data are higly right skewed.

# Descriptive and Inferential Statistics

## Provide univariate descriptive statistics and appropriate plots for the training data set. 

```{r}
summary(df)
```


```{r fig.height=10, fig.width=12}
df %>% 
  select_if(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(x=value)) +
  geom_histogram() +
  facet_wrap(~key, ncol = 4, scales = 'free') +
  theme_minimal()
```


```{r fig.height=10, fig.width=12}
df %>% 
  select_if(is.factor) %>% 
  gather() %>% 
  ggplot(aes(x=value)) +
  geom_bar(stat = 'count') +
  facet_wrap(~key, ncol = 4, scales = 'free') +
  theme_minimal()
```


## Provide a scatterplot of X and Y. 

```{r }
ggplot(df, aes(LotArea, SalePrice))+
  geom_point(colour = 'skyblue', alpha = 0.3) +
  theme_minimal()
```


```{r fig.height=10, fig.width=12}
df %>% 
  select_if(is.numeric) %>% 
  gather(VariableName, Value, -SalePrice, -Id) %>% 
  ggplot(aes(SalePrice,Value))+
  geom_point(colour = 'gray28', alpha = 0.3) +
  facet_wrap(~VariableName, ncol = 4, scales = 'free') +
  theme_minimal()
```

```{r fig.height=10, fig.width=12}
df %>% 
  dplyr::select(which(sapply(.,class)=="factor"), SalePrice) %>% 
  gather(VariableName, Value, -SalePrice) %>%
  ggplot(aes(SalePrice,Value))+
  geom_point(colour = 'gray28', alpha = 0.3) +
  facet_wrap(~VariableName, ncol = 4, scales = 'free') +
  theme_minimal()
```


## Derive a correlation matrix for any THREE quantitative variables in the dataset. 

```{r}
vars <- c('LotArea', 'SalePrice', 'GrLivArea')

corDF <- df %>% dplyr::select(LotArea, SalePrice, GrLivArea) %>% cor(use = "complete.obs")
corDF
library(corrplot)
corrplot(corDF)
```




## Test the hypotheses that the correlations between each pairwise set of variables is 0 and provide a 92% confidence interval. 



```{r}
testCor <- df %>% dplyr::select(LotArea, SalePrice, GrLivArea)
cor.test(testCor$LotArea, testCor$SalePrice, conf.level = 0.92)
cor.test(testCor$LotArea, testCor$GrLivArea, conf.level = 0.92)
cor.test(testCor$SalePrice, testCor$GrLivArea, conf.level = 0.92)
```

## Discuss the meaning of your analysis. 

All the correlations were significant.

*Would you be worried about familywise error? Why or why not?*

Familywise error is applicable when making multiple hypothesis tests. 

$$
\alpha = 1 - (1-p)^c
$$
Where $p$ is the confidence level and $c$ is the number of tests. 

```{r}
Alpha <- 1 -  0.92^3
Alpha
```

So we have about a 22% chance of making a type one error. This is definitely something to be worried about, hwoever, given that the p values were below 0.08, we have less to worry about.



# Linear Algebra and Correlation.

## Invert your 3 x 3 correlation matrix from above. 

(This is known as the precision matrix and contains variance inflation factors on the diagonal.) 


```{r}
corDFT <- t(corDF)
corDFT
```


## Multiply 
the correlation matrix by the precision matrix, and then multiply the precision matrix by the correlation matrix. 

```{r}
corDF %*% corDFT
corDFT %*% corDF
```


## Conduct LU decomposition on the matrix.

```{r}
library(pracma)
luMat <- corDFT %*% corDF
lu(luMat, scheme = "ijk")
```

# Calculus-Based Probability & Statistics

Many times, it makes sense to fit a closed form distribution to data. For the first variable that you selected which is skewed to the right, shift it so that the minimum value is above zero as necessary.

```{r}
min(df$LotArea) # Above zero
lotVec <- df$LotArea - min(df$LotArea) + 0.0001
min(lotVec)
```


Then load the MASS package and run fitdistr to fit an exponential probability density function. (See https://stat.ethz.ch/R-manual/R-
devel/library/MASS/html/fitdistr.html ). 

```{r}
library(MASS)
```

Find the optimal value of λ for this distribution, and then take 1000 samples from this exponential distribution using this value (e.g., rexp(1000, λ)). 

```{r}
fit <- fitdistr(lotVec, "exponential")
lambda <- fit$estimate
lambda
samples <- rexp(1000, lambda)
```


Plot a histogram and compare it with a histogram of your original variable. 
```{r}
qplot(samples, main = 'Simulated') + theme_minimal()
qplot(lotVec, main = 'Observed') + theme_minimal()
```


Using the exponential pdf, find the 5th and 95th percentiles using the cumulative distribution function (CDF). 

$$
CDF = \frac{ln(1 - P)}{-\lambda}
$$

Also generate a 95% confidence interval from the empirical data, assuming normality. Finally, provide the empirical 5th percentile and 95th percentile of the data. Discuss.

```{r}
cdf5 <- log(0.95)/ - lambda
cdf95 <- log(0.05)/ - lambda

obs5 <- quantile(lotVec, 0.05)
obs95 <- quantile(lotVec, 0.95)

x1 <- t.test(lotVec)$conf.int[1:2]

x <- data_frame(Type = c('Simulated', "Observed", "CI"), P5 = c(cdf5, obs5, x1[1]), P95 = c(cdf95, obs95, x1[2])) 
kable(x)


```

The exponential distroubution does not do a good job of estimating the actual variables. In this case, a boot straped CI would proabably be more effective. The Actual t-test of the mean doesn't have much to do with the quantiles. 

# Modeling

Build some type of multiple regression model and submit your model to the competition board. Provide your complete model summary and results with analysis. Report your Kaggle.com user name and score.



## Data cleaning:




```{r}
cleaner <- function(df){
  numericNames <- df %>% select_if(is.numeric) %>% colnames()
  for (x in numericNames) {
     meanValue <- mean(df[[x]],na.rm = TRUE)
      df[[x]][is.na(df[[x]])] <- meanValue
  }
  # Skewed fix
  for (i in numericNames) {
    if (skewness(df[[i]]) > 0.75){
      df[[i]] = log(df[[i]] + 1)
    }
  }
  df[is.na(df)] <- 'Unknown'
  return(df)
}
trainDF <- read_csv('train.csv')
testDF <- read_csv('test.csv')
trainDF$IsTrain <- TRUE
testDF$IsTrain <- FALSE
SalePrice <- log(trainDF$SalePrice)
trainDF$SalePrice <- NULL
Id <- testDF$Id
aggDF <- rbind(trainDF, testDF)
aggDF$Id <- NULL
aggDF <- cleaner(aggDF)

aggDF[sapply(aggDF, is.character)] <- lapply(aggDF[sapply(aggDF, is.character)], as.factor)
aggDF[sapply(aggDF, is.factor)] <- lapply(aggDF[sapply(aggDF, is.factor)], as.numeric) # possible because of non-linear fitting

trainDF <- aggDF %>% filter(IsTrain == TRUE)  %>% dplyr::select(-IsTrain)
trainDF$SalePrice <- SalePrice
testDF <- aggDF %>% filter(IsTrain == FALSE) %>% dplyr::select(-IsTrain)


```


```{r}
svmMod <- svm(SalePrice ~ ., data = trainDF, cost = 1)
preds = predict(svmMod, testDF)
submission = cbind(Id = Id, SalePrice = exp(preds))
colnames(submission) = c("Id","SalePrice")

write.csv(submission,file="svmKai.csv",row.names=FALSE)
# Kaggle Score: 0.16514

```

```{r}
predsActual = predict(svmMod, trainDF)
qplot(x = exp(predsActual), 
      y =  exp(trainDF$SalePrice), 
      xlab = 'Actual Sale Price',
      ylab = 'Predicted Sale Price',
      main = 'Scatter Plot') + 
  theme_minimal()
```


```{r}
train <- read.csv("train.csv")


linearMod <-lm(SalePrice ~ YearBuilt, data=train)


linPred <- predict(linearMod, train)
qplot(train$SalePrice, 
      linPred,
      xlab = 'Actual',
      ylab = 'Predicted',
      main = 'Scatter for Linear Regression') +
  theme_minimal()

# 0.33 on the training set.

```

We can see this looks very non-linear. This is why the SVM worked so much better. 


```{r}
max(abs(trainDF$SalePrice - predsActual))
diagnostic <- trainDF
diagnostic$Preds <- predsActual

diagnostic %>% 
  filter(abs(SalePrice - Preds) > 0.3 )
```


* Kaggle User Name: https://www.kaggle.com/kailukowiak
* Kaggle Score: 0.16514



