---
title: "Assignment 2"
author: "Kai Lukowiak"
date: '2018-02-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Problem Set 1
## 1. Show that $A^T A \neq AA^T$ in general. (Proof and demonstration.)

Proof by contradiction:
Let:
$$
A =
\begin{bmatrix}
    a & b \\
    c & d \\
\end{bmatrix}
\\
A^T = 
\begin{bmatrix}
    a & c \\
    b & d \\
\end{bmatrix}
$$

Therefore:

$$
A^TA = 
\begin{bmatrix}
    a^2+b^2 & ac + bd \\
    ac +bd & c^2 + d^2 \\
\end{bmatrix}
\\
AA^T = 
\begin{bmatrix}
    a^2+c^2 & ab + cd \\
    ab - bd & b^2 + d^2 \\
\end{bmatrix}
$$

Thus, we can see that $\forall\quad b \neq c$ the product of $A^TA \neq AA^T$. Therefor in general, unless the transposes are equal, they will not produce the same result. 

```{r}
A <- matrix(c(1,3,2,4), ncol = 2)

identical(t(A) %*% A, A %*% t(A))
```

So we can see they are not identical, however...

```{r}
A <- matrix(c(1,3,3,4), ncol = 2)
identical(t(A) %*% A, A %*% t(A))
```
So when the matrix can be inverted, it is true. 

## 2. 
For a special type of square matrix A, we get $A^T A = AA^T$. Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix).

See extended answer to 1. In general if the matrix has a transpose equal to the original matrix it will be have this characteristic. 

# Problem Set 2
Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorization. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars. Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer. 


Unfortunately the solution we used for last weeks assignment is a bit complex to use for this case. Instead we will you a more basic function that only worries about creating an upper triangular matrix and then take the negative of the row multiplier and insert it into the correct position.



```{r}
set.seed(605)
n <- 5
A <- matrix(sample.int(10, size = n^2, replace = TRUE), ncol = n)
A
```


```{r}
matrixLU <- function(A){
  U = A # to keep consistant formatting.
  L = diag(x = 1, ncol = ncol(A), nrow = nrow(A)) # Square diagonal to populate cells.
  
  # For loops to loop over elements in L and U
  for (i in 1:(nrow(U) - 1)){
    for (j in (i + 1):nrow(U)){
      if (U[i, i] != 0){
        multiplier = U[j, i] / U[i, i]
        L[j, i] = multiplier
        U[j, ] = U[j, ] - multiplier * U[i, ]
      }
    }
  }
  return(list('L' = L, 'U' = U))
}

X <- matrixLU(A)
X$L
X$U
X$L %*% X$U
```


```{r}
B <-  matrix(c(1:16), ncol = 4)
X1 <- matrixLU(B)
X1$L
X1$U
all.equal(X1$L %*% X1$U, B)
```

