---
title: "Assignment 2"
author: "Kai Lukowiak"
date: '2018-02-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Problem Set 1
## 1. Show that $A^T A \neq AA^T$ in general. (Proof and demonstration.)

Proof by contradiction:
Let:
$$
A =
\begin{bmatrix}
    a & b \\
    c & d \\
\end{bmatrix}
\\
A^T = 
\begin{bmatrix}
    a & c \\
    b & d \\
\end{bmatrix}
$$

Therfore:

$$
A^TA = 
\begin{bmatrix}
    a^2+b^2 & ac + bd \\
    ac +bd & c^2 + d^2 \\
\end{bmatrix}
\\
AA^T = 
\begin{bmatrix}
    a^2+c^2 & ab + cd \\
    ab - bd & b^2 + d^2 \\
\end{bmatrix}
$$

Thus, we can see that $\forall\quad b \neq c$ the product of $A^TA \neq AA^T$. Therefor in general, unless the transposes are equal, they will not produce the same result. 


## 2. 
For a special type of square matrix A, we get $A^T A = AA^T$. Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix).

See extended answer to 1. In general 

# Problem Set 2
Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars. Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer. 


Unfortunetly the solution we used for last weeks assignment is a bit complex to use for this case. Instead we will you a more baisic funciton that only worries about creating an upper trianglular matrix and then take the negative of the row multiplier and insert it into the correct position.



```{r}
set.seed(605)
n <- 5
A <- matrix(sample.int(10, size = n^2, replace = TRUE), ncol = n)
A
```


```{r}
matrixLU <- function(A){
  U = A # to keep consistant formatting.
  L = diag(x = 1, ncol = ncol(A), nrow = nrow(A)) # Square diagonal to populate cells.
  
  
  
  return(list('U' = U, 'L' = L))
}

matrixLU(A)
```

